---
title: "code"
author: "melonheader"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

##########
## INFO ##
##########
This R notebook contains the source code for the plots that are used in 
the manuscript from 03.04.2023


```{r}
## HEADER
set.seed(42)

library(tidyverse)
library(GGally)
library(Guitar)
library(tictoc)
library(ggridges)
library(DESeq2)
library(purrr)
library(janitor)
library(patchwork)
library(magrittr)
library(gprofiler2)

## auxiliary
source("F_HelpersMetaprofile.R")
source("F_HelpersRNAseq.R")
source("F_HelpersDistMod.R")
source("F_miscellaneous.R")
## loc pal 3 way
pal_loc.3way <- c(Projections = "#82b21e", Cytoplasm = "#006beb", 
                  NS = "gray50", Nucleus = "#236477")
## loc pal 2 way
pal_loc.2way <- c(Neurite = "#82b21e", NS = "gray50", Soma = "#006beb")

##
pal_mut <- c(GFP = "steelblue3", NS = "gray50", dnCaf1 = "orangered3")
##
pal_mut.msa <- c("#f18ea3", "grey50", "#481f31")
##
pal_msa <- c("#009cb8", "grey50", "#d64ecf")


mm10Ann <- read_tsv("../data/mm10_annotation.tsv", show_col_types = F)


##
ggplot2::theme_set(theme_bw(base_size = 14) +
                     theme(panel.border = element_blank(),
                           axis.line = element_line(size = 0.25, color = "black"), 
                           axis.ticks = element_line(size = 0.25, color = "black"),
                           axis.text = element_text(color = "black"),
                           legend.background = element_blank()
                     )
)
```



########################################################
#   1. m6A related plots  
################################################################################
## 1.1 Grid of correlation plots. 
Log2FoldChagnes from m6A-RIPseqs are plotted against each other.
Either IP vs Flow or IP vs Input are used. Right-upper triangle of the plot grid shows Pearson correlations between the respective datasets in the left-lower triangle.
```{r}
de_m6a <- read_rds("../data/m6ARIP.lg2fc_list.RDS")
de_m6a_plot <- purrr::map2(de_m6a, 
                           names(de_m6a),
                           ~ .x %>% 
                             filter(log2(baseMean) > 5.5) %>% 
                             dplyr::select(gene, log2FoldChange) %>% 
                             `colnames<-`(c("gene", paste0(.y)))) %>% 
  purrr::reduce(., inner_join, by = "gene") %>% 
  `[`(, c("gene", 
          "soft-filt_outfilt_IP_unfr_m6a.sysy_vs_FLOW_unfr_m6a.sysy",
          "soft-filt_outfilt_IP_unfr_m6a.abcam_vs_FLOW_unfr_m6a.abcam", 
          "soft-filt_IP_fr_m6a.sysy_vs_INPUT_fr_total", 
          "soft-filt_IP_fr_m6a.abcam_vs_INPUT_fr_total", 
          "soft-filt_ip_synaptosome_vs_input_synaptosome",
          "soft-filt_ip_homogenized_forebrain_vs_input_homogenized_forebrain",
          "soft-filt_outfilt_m6Am_RIP_WT_vs_input_RNA_WT"
          )
      ) %>%  
  `colnames<-`(c("gene", 
                 "A", 
                 "B", 
                 "C", 
                 "D",  
                 "E", 
                 "F", 
                 "G")
               )

# Check correlations (as scatterplots), distribution and print corrleation coefficient 
my_point <- function(data, mapping, ..., low = "#132B43", high = "#56B1F7") {
  ggplot(data = data, mapping = mapping) +
    #geom_point(...) +
    geom_bin2d(binwidth = 0.5, ...) +
    #geom_smooth(y ~ x, method = "lm", se = F) +
    scale_fill_viridis_c() +
    #scale_fill_gradient(low = low, high = high) + 
    scale_x_continuous(breaks = c(-2.5, 0, 2.5), limits = c(-5, 5)) +
    scale_y_continuous(breaks = c(-2.5, 0, 2.5), limits = c(-5, 5))
}

ggpairs(de_m6a_plot[, -1], 
        title = "m6A datasets comparison",
        legend = c(2, 1),
        lower = list( 
         continuous = my_point
         ), 
        progress = FALSE) -> p.1.1


ggsave(filename = "../output/p.1.1.pdf", p.1.1)
```


################################################################################
## 1.2 Grid showing intersections between m6A-peaks called from different m6ARIP-seqs datasets.
Left-upper triangle of the grid shows the number of intersecting peaks between datasets on X agaisnt datasets on Y.
Right-lower triangle shows the opposite, i.e. datasets on Y against X.
```{r}
intersect_files <- list.files("../data/PeakCalls/Intersections/", 
                              pattern = "inter", 
                              full.names = T) %>% 
  grep("-KO", ., value = T, invert = T)

# create alphabet for msa datasets
alphbet_msa <- tibble(id_msa = c("3", "4", "5", "6", "7"), 
                      dataset_msa = c("m6A_Sysy",
                                      "m6A_Abcam", 
                                      "Merkurjev_Synaptosome", 
                                      "Merkurjev_Homogenized-brain", 
                                      "Engel_WT"))

names(intersect_files) <- basename(intersect_files) %>% 
  str_remove_all(., ".bed_intersect_out.bed") %>% 
  enframe() %>% 
  left_join(., alphbet_msa, by = c("value" = "dataset_msa")) %>% 
  pull(id_msa)

int_dat <- purrr::map2(intersect_files, 
                       names(intersect_files), 
                       ~ read_tsv(.x, col_names = F, show_col_types = F) %>% 
                         group_by(X13) %>% 
                         summarise(int_count = n()) %>% 
                         ungroup() %>% 
                         dplyr::mutate(query_bed = .y, 
                                       X13 = str_remove_all(X13, ".bed")) %>% 
                         dplyr::rename(base_bed = X13)) %>% 
  purrr::reduce(., rbind) %>% 
  filter(!base_bed %in% c("Engel_FTO-KO", "Engel_Mettl-KO")) %>%
  inner_join(., 
             alphbet_msa, 
             by = c("base_bed" = "dataset_msa")
             ) %>% 
  mutate(base_bed = factor(id_msa, levels = c("3", "4", "5", "6", "7")),
         query_bed = factor(query_bed, levels = c("3", "4", "5", "6", "7")), 
				 color_var = ifelse(query_bed != id_msa, "comp", "same")) 

##
ggplot(data = int_dat, 
       aes(x = query_bed, 
           y = base_bed, 
           fill = int_count)) + 
  geom_tile(aes(color = color_var), 
            linejoin = "bevel",
            width=.9, height=.9, 
            size = 0.5) + 
  scale_color_manual(values = c(adjustcolor("#333333", alpha.f = 0.1), "red")) +
  scale_fill_gradient(low = "#e5ebf0", 
                      high = "#325f87") +
  geom_text(aes(label = int_count)) +
  theme(axis.text.x = element_text(angle = 0), axis.text.y = element_text(angle = 0)) +
  labs(fill = "Peak intersection, n", y = "", x = "") + 
  guides(color = "none") -> p.1.2

ggsave(filename = "../output/p.1.2.pdf", p.1.2)
```


################################################################################
## 1.3 MA-plots for total(unfragmented) m6ARIPseq datasets.
Mean abundance is plotted against lg2FC between eluate and flowthrough.
Each plot corresponds to separate antibody that are labeled within the code.
```{r}
## Abcam antibody
plot_ma(de_res = de_m6a$`soft-filt_outfilt_IP_unfr_m6a.abcam_vs_FLOW_unfr_m6a.abcam`, 
        lg2cts_cutoff = 5.5, lg2fc_cutoff = 1, padj_cutoff = 0.05, y_cutoff = c(-10, 10),
        custom_palette = pal_msa[c(1, 3, 2)], empty = F) -> p.1.3.abcam
## Sysy antibody
plot_ma(de_res = de_m6a$`soft-filt_outfilt_IP_unfr_m6a.sysy_vs_FLOW_unfr_m6a.sysy`, 
        lg2cts_cutoff = 5.5, lg2fc_cutoff = 1, padj_cutoff = 0.05, y_cutoff = c(-10, 10),
        custom_palette = pal_msa[c(1, 3, 2)], empty = F) -> p.1.3.sysy

purrr::map2(c("p.1.3.abcam", "p.1.3.sysy"),
            list(p.1.3.abcam, p.1.3.sysy),
            ~ ggsave(filename = paste0("../output/", .x, ".pdf"), .y))
```


################################################################################
## 1.4 Metagene-profile plot.
Shows the distribution of the detected m6A peaks along the features of a metagene. 
 
 !Alert: access to the GTF annotation that was used for the original plot is currently restricted on the server. 
 Same version of the annotation that was reported in the initial code was downloaded from Ensembl.
 
 !The chunk below prepares the genomic annotation for metaprofile plotting. In the distribution, this step is already done and the annonation can be loaded from ../data. If the user wishes to rerun this step, he needs to run gunzip .data/Mus_musculus.GRCm38.96.gtf.gz in the root folder to unzip the gtf file.
```{r, eval=F}
# import different data formats into a named list object.
# These genomic features are using mm10 genome assembly
peak_dir <- "../data/PeakCalls/m6a_fragmented"
peak_sites <- list(file.path(peak_dir, "m6a_fragmented_IP_fr_m6a-abcam", "con_peak.bed"), 
                   file.path(peak_dir, "m6a_fragmented_IP_fr_m6a-sysy", "con_peak.bed"))
tx_gtf <- "../data/Mus_musculus.GRCm38.96.gtf" 

# Get table for metprofile plotting
tic()
metap_list <- get_metaprofile_tab(txGTF = tx_gtf,
                                  stBedFiles = peak_sites,
                                  headOrtail = TRUE,
                                  enableCI = FALSE,
                                  mapFilterTranscript = TRUE,
                                  pltTxType = c("mrna"),
                                  stGroupName = c("Abcam", "Sysy"))
toc()

## save data for simpler re-usage
write_rds(metap_list, "../data/p.1.4_metaplist.RDS")
```

```{r}
## The data is preprocessed; load the prepared list; To assemble the genomic coordinates again, re-run the previous chunk
metap_list <- read_rds("../data/p.1.4_metaplist.RDS")
### assemble the plot
metap_tab <- metap_list$data_tab
# c("#592f93", "#f8a51b")
ggobj <- ggplot(data = metap_tab, aes(x = metap_tab %>% pull(x))) +
  geom_line(aes(y = density, 
                color = group), 
            alpha = 1,
            size = 1) +
  geom_ribbon(aes(ymin = rep(0, length(density)),
                  ymax = density, 
                  fill = group), 
              alpha = 0.2) + 
  #theme_metaprofile(aspect.ratio = 0.6, base_size = 16) + 
  xlab("") + labs(fill = "Antibody:", color = "Antibody:") +
  scale_color_manual(values = c("#999999", "#333333")) + 
  scale_fill_manual(values = c("#999999", "#333333"))
#
pos <- Guitar:::.generate_pos_para(4)
p.1.4 <- Guitar:::.RNAPlotStructure(p = ggobj, componentWidth = metap_list$component_width, headOrtail = TRUE, pos = pos)
#
ggsave(filename = "../output/p.1.4.pdf", p.1.4)
```


################################################################################
## 1.5 Plots showing distribution of a DRACH motif around the peak summits inferred from m6A-RIPseqs
```{r, fig.width=8, fig.height=5}
# Distribution of motif relative to peak summit
m_dist_sy <- quiet(read_tsv("../data/PeakCalls/m6a_fragmented/m6a_fragmented_IP_fr_m6a-sysy/homer_hist_200.tsv", 
                         col_types = cols(.default = col_double()))
                ) %>% 
  dplyr::slice(2:(length(.[[1]]) - 1)) %>% 
  `colnames<-`(c("bp", "motif_total", "motif_sense", "motif_antisense", paste0(c("A", "C", "G", "T"), "_freq"))) %>% 
  dplyr::select(-contains("freq"), -motif_total) %>% 
  gather(., contains("motif"), key = "motif", value = "value") %>% 
  mutate(antibody = "Sysy")
m_dist_ab <- quiet(read_tsv("../data/PeakCalls/m6a_fragmented/m6a_fragmented_IP_fr_m6a-abcam/homer_hist_200.tsv", 
                         col_types = cols(.default = col_double()))
                ) %>% 
  dplyr::slice(2:(length(.[[1]]) - 1)) %>% 
  `colnames<-`(c("bp", "motif_total", "motif_sense", "motif_antisense", paste0(c("A", "C", "G", "T"), "_freq"))) %>% 
  dplyr::select(-contains("freq"), -motif_total) %>% 
  gather(., contains("motif"), key = "motif", value = "value") %>% 
  mutate(antibody = "Abcam")
m_dist <- rbind(m_dist_sy, m_dist_ab)

  
ggplot(data = m_dist,
       aes(x = bp, 
           y = value, 
           color = motif)) + 
  geom_point(size = 2) + 
  geom_path(size = 1) + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_manual(values = alpha(c(pal_msa[3], "gray60"), 0.67), 
                     name = "Strand:", 
                     breaks = c("motif_sense", "motif_antisense"), 
                     labels = c("Peak strand", "Opposite strand")) +
  labs(x = "Distance from the peak summit, bp", 
       y = "Motif count per 5 bp, per peak") + 
  facet_grid(. ~ antibody) +
  theme(aspect.ratio = 1.25) -> p.1.5

ggsave("../output/p.1.5.pdf", p.1.5)
```


################################################################################
## 1.6 2-dimensional histogram showing the correlation between different antibodies used for total(unfragmented) m6A-RIPseqs.
log2FC between Eluate and Flowthrough estimated using SySy antibody is plotted against lg2FC between Eluate and Flowthrough estimated using Abcam antibody. Pearson correlation between X and Y is reported on the plot.
```{r}
abcam_tab <- de_m6a$`soft-filt_outfilt_IP_unfr_m6a.abcam_vs_FLOW_unfr_m6a.abcam` %>% dplyr::select(gene, lgfc_abcam = log2FoldChange)
sysy_tab <- de_m6a$`soft-filt_outfilt_IP_unfr_m6a.sysy_vs_FLOW_unfr_m6a.sysy` %>% dplyr::select(gene, lgfc_sysy = log2FoldChange)
ab_cor <- inner_join(abcam_tab, sysy_tab, by = "gene")
#
ggplot(data = ab_cor, 
       aes(x = lgfc_sysy, 
           y = lgfc_abcam)) + 
  geom_bin2d(binwidth = 0.25) + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "#333333") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#333333") +
  geom_smooth(method = "lm", se = F) +
  ggpubr::stat_cor() + 
  scale_fill_viridis_c() + 
  theme_bw(base_size = 14) + 
  lims(x = c(-10, 10), y = c(-10, 10)) +
  labs(y = "log2FC IP vs Flow (abcam)", x = "log2FC IP vs Flow (sysy)") -> p.1.6

ggsave("../output/p.1.6.pdf", p.1.6)
```

################################################################################
## 1.7 qPCR validation of m6ARIPseq experiments.
Pearson correlations between lg2FCs from m6A-RIPseqs (Y) and delta Ct's for candidate transcripts (X) are reported on the plot for both antibodies.
```{r}
f_tab <- tibble(dct_abcam = c(5.81755383365437,
                              5.54591915031877,
                              4.57699239733848,
                              3.59975026782392, 
                              2.21033725804462, 
                              1.58741227582733, 
                              1.55375708209785, 
                              1.23728134509787, 
                              1.43987527422602), 
                      dct_sysy = c(6.78873338743332, 
                                   5.79912705284167, 
                                   7.44078689636392, 
                                   7.63113116282434, 
                                   0.776246183340399, 
                                   2.50182008783007, 
                                   1.21739340149005, 
                                   0.414522585525116, 
                                   -0.655463079309084), 
                      meth_group = c(rep("high_methylation", 4), 
                                     rep("low_methylation", 5)),
                      gene_name = c("Cdr2", 
                                    "Sh3gl1", 
                                    "Ngfr", 
                                    "Sacs", 
                                    "Gapdh", 
                                    "Actb",
                                    "Igf2", 
                                    "Eif5a", 
                                    "Ndufa4"), 
                      gene_id = c("ENSMUSG00000030878", 
                                  "ENSMUSG00000003200", 
                                  "ENSMUSG00000000120", 
                                  "ENSMUSG00000048279", 
                                  "ENSMUSG00000057666", 
                                  "ENSMUSG00000029580", 
                                  "ENSMUSG00000048583", 
                                  "ENSMUSG00000078812", 
                                  "ENSMUSG00000029632")
                      )
de_abcam <- de_m6a$`soft-filt_outfilt_IP_unfr_m6a.abcam_vs_FLOW_unfr_m6a.abcam` %>% 
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1], 
         gene_name = str_split_fixed(gene, "_", 2)[, 2]) %>% 
  dplyr::select(gene_id, gene_name, l2fc_abcam = log2FoldChange)
de_sysy <- de_m6a$`soft-filt_outfilt_IP_unfr_m6a.sysy_vs_FLOW_unfr_m6a.sysy` %>% 
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1], 
         gene_name = str_split_fixed(gene, "_", 2)[, 2]) %>% 
  dplyr::select(gene_id, gene_name, l2fc_sysy = log2FoldChange)

f_tab <- left_join(f_tab, de_abcam, by = c("gene_id", "gene_name")) %>% 
  left_join(., de_sysy, by = c("gene_id", "gene_name")) #%>% 

ft1 <- f_tab %>% 
  dplyr::select(gene_name, gene_id, l2fc_abcam, l2fc_sysy) %>% 
  gather(., contains("l2fc"), key = "ab", value = "lg2fc") %>% 
  mutate(ab = str_remove_all(ab, "l2fc_"))

f_tab <- f_tab %>% 
  dplyr::select(gene_name, gene_id, meth_group, dct_abcam, dct_sysy) %>% 
  gather(., contains("dct"), key = "ab", value = "dct") %>%
  mutate(ab = str_remove_all(ab, "dct_")) %>%
  inner_join(., ft1, by = c("gene_id", "gene_name", "ab"))

ggplot(data = f_tab, 
       aes(x = dct, 
           y = lg2fc, 
           shape = ab, 
           color = ab)) + 
  geom_point(size = 2.5) + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
  ggpubr::stat_cor() + 
  ggrepel::geom_text_repel(aes(label = gene_name), min.segment.length = 0.7, force = 100) + 
  #scico::scale_color_scico_d(palette = "batlow") +
  scale_color_manual(values = c("#999999", "#333333")) +
  theme_bw(base_size = 14) + 
  labs(y = "log2FoldChange", x = "deltaCT") -> p.1.7

ggsave("../output/p.1.7.pdf", p.1.7)
```



################################################################################
## 1.8 Validation of m6A-RIP protocol by RT-qPCR; 
```{r, fig.width=5, fig.height=6}
e_tab <- tibble(group = c("Input", 
                          "Input",
                          "Input",
                          "Input", 
                          "Flow_m6A", 
                          "Flow_m6A", 
                          "Flow_m6A", 
                          "Flow_m6A",
                          "Flow_m6A", 
                          "Flow_m6A", 
                          "IP_m6A", 
                          "IP_m6A", 
                          "IP_m6A", 
                          "IP_m6A",
                          "IP_m6A", 
                          "IP_m6A"), 
                rep = c(1, 2, 1, 2, 
                        1, 2, 1, 2,
                        1, 2, 
                        1, 2, 1, 2,
                        1, 2),
                rluc_group = c("m6A Rluc",
                               "m6A Rluc", 
                               "Rluc",
                               "Rluc", 
                               "m6A Rluc",
                               "m6A Rluc", 
                               "Rluc",
                               "Rluc",  
                               "m6A Rluc (beads only)", 
                               "m6A Rluc (beads only)",
                               "m6A Rluc",
                               "m6A Rluc", 
                               "Rluc",
                               "Rluc",  
                               "m6A Rluc (beads only)", 
                               "m6A Rluc (beads only)"),
                rluc_value = c(104.36, 
                               95.64,
                               95.33,
                               97.23,
                               1.20,
                               1.28,
                               69.00,
                               101.57, 
                               101.57,
                               94.99,
                               35.94,
                               40.41,
                               0.79,
                               1.25,
                               0.05,
                               0.04)
                ) %>% 
  mutate(group = factor(group, levels = c("Input", 
                                          "Flow_m6A",
                                          "Flow_beads", 
                                          "IP_m6A", 
                                          "IP_beads"))) %>% 
  group_by(group, rluc_group) %>%
  summarise(rluc_mean = mean(rluc_value), 
            S = var(rluc_value), 
            stddev = sd(rluc_value),
            stderr = qnorm(0.975) * (stddev / sqrt(n()))
            )

#
ggplot(data = e_tab,
       aes(x = group,
           y = rluc_mean, 
           fill = rluc_group
           )
       ) + 
  geom_col(width = 0.8, 
           alpha = 0.67,
           position = position_dodge(width = 0.9), 
           color = "#333333"
           ) + 
  geom_errorbar(aes(ymin = rluc_mean - stddev, 
                    ymax = rluc_mean + stddev),
                position = position_dodge(width=0.9), 
                width = 0.25) +
  theme(aspect.ratio = 2) +
  scale_y_continuous(name = "% of transcript compared to input", 
                     breaks = c(0, 25, 50, 75, 100, 125)
                     ) +
  labs(x = "") + 
  #scico::scale_colour_scico_d(palette = "lajolla") +
  scale_fill_manual(values = c("#d64ecf","#333333", "#009cb8")) -> p.1.8

ggsave("../output/p.1.8.pdf", p.1.8)
```

################################################################################
## 1.9 Scatter plot showing the difference between Neuritic and Somatic half-lives of m6A-enriched and m6A-depleted transcripts
```{r}
data_loc.Ascl1Slam <- read_tsv("../data/Ascl1.Compartments_Slamseq/soft-filt_outfilt_N_0_vs_S_0_deseq_res.tsv")
read_tsv("../data/hard-filt_outfilt_IP_unfr_m6a.sysy_vs_FLOW_unfr_m6a.sysy_deseq_res.tsv", 
         show_col_types = F) %>% 
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
         gene_name = str_split_fixed(gene, "_", 2)[, 2]) %>% 
  dplyr::select(-gene) -> res_msa.sysy
##
t.hl_comp.1 <- read_tsv("../data/Ascl1.Compartments_Slamseq/data_hl.tsv")
t.hl_comp.1 %>% 
  pivot_wider(., 
              names_from = "Group",
              values_from = c("Half_life", "HL_LoInt95", "HL_UpInt95"), 
              names_glue = "{Group}_{.value}") -> t.hl_scat
t.hl_scat.filt <- t.hl_scat %>% 
  filter((Soma_HL_LoInt95 - Soma_HL_LoInt95) < Soma_Half_life,
         (Neurites_HL_UpInt95 - Neurites_HL_LoInt95) < Neurites_Half_life)

res_msa.sysy %>% 
  filter(lfcSE < quantile(lfcSE, 0.9)) %>% 
  mutate(sign_msa = case_when(
           log2FoldChange > 1 & padj < 0.05 ~ "m6A-enriched",
           log2FoldChange < -1 & padj < 0.05 ~ "m6A-depleted",
           T ~ "NS"
         )) %>% 
  dplyr::select(gene_id, gene_name, sign_msa) %>% 
  inner_join(data_loc.Ascl1Slam %>% 
               mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
                      gene_name = str_split_fixed(gene, "_", 2)[, 2]) %>% 
               dplyr::select(-gene), 
             ., 
             by = c("gene_id", "gene_name")) %>% 
  mutate(sign_loc = case_when(
           log2FoldChange > log2(1.5) & padj < 0.05 ~ "Neurite",
           log2FoldChange < -log2(1.5) & padj < 0.05 ~ "Soma",
           T ~ "NS"
         )) %>% 
  dplyr::select(contains("gene"), contains("sign"), everything()) %>% 
  inner_join(t.hl_scat.filt %>% mutate(delta_hl = Neurites_Half_life - Soma_Half_life), 
             ., 
             by = c("gene_id", "gene_name")) -> plchld
  
plchld %>% 
  filter(sign_loc != "NS",
         sign_msa != "NS") %>% 
  ggplot(data = ., aes(x = Soma_Half_life, y = Neurites_Half_life)) + 
  geom_point(alpha = 0.3, aes(color = sign_msa)) +
  geom_abline(intercept = 0, slope = 1) +
  scale_color_manual(values = pal_msa[c(1,3)]) +
  scale_x_continuous(breaks = seq(0, 16, by = 4), limits = c(0, 16)) +
  scale_y_continuous(breaks = seq(0, 16, by = 4), limits = c(0, 16)) +
  facet_grid(. ~ sign_loc) +
  theme(aspect.ratio = 1) -> p.1.9
ggsave("../output/p.1.9.pdf", p.1.9)

l.pvsl <- purrr::map(split(p.1.9$data, p.1.9$data$sign_loc), ~ wilcox.test((Neurites_Half_life - Soma_Half_life) ~ sign_msa, data = .x)$p.value)
tibble(compartment = names(l.pvsl), test = "Wilcoxon", p.value = purrr::reduce(l.pvsl, c)) %>% mutate(p.adj = p.adjust(p.value, "fdr")) -> p.1.9_stat
write_tsv(p.1.9_stat, "../output/p.1.9_stats.tsv")
```

```{r}
plchld %>% 
  filter(sign_loc != "NS",
         sign_msa != "NS") %>% 
  group_by(sign_loc) %>% 
  nest() %>% 
  mutate(test = "Wilcoxon",
         p.value = purrr::map_dbl(data,
                                  ~ wilcox.test(formula = delta_hl ~ sign_msa, data = .x)$p.value)
         ) %>% 
  ungroup() %>% 
  dplyr::select(-data) %>% 
  mutate(p.adj = p.adjust(p.value)) %>% 
  write_tsv(., "../output/p.1.9_stats.tsv")
```



########################################################  
#   2. Plots related to Figure 1
################################################################################
## 2.1 Density plots showing the distributions of Half-lives in subcellular compartments
```{r}
l.hl.comps <- read_rds("../data/CompsHLs_list.RDS")
rbind(
  l.hl.comps$PCN.3c %>%
    dplyr::select(-contains("95")) %>%
    mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
           gene_name = str_split_fixed(gene, "_", 2)[, 2],,
           group = case_when(
             group == "Pro" ~ "Neurite, PCN",
             group == "Nuc" ~ "Nucleus, PCN",
             group == "Cyto" ~ "Soma, PCN",
             T ~ "ayy"
           )) %>%
    dplyr::select(gene_id, gene_name, group, Half_life),
  l.hl.comps$Ascl1.2c %>%
    dplyr::select(-contains("95")) %>%
    mutate(group = case_when(
             Compartment == "Neurites" ~ "Neurite, mESC",
             Compartment == "Soma" ~ "Soma, mESC",
             T ~ "ayy"
           )) %>% 
    filter(Half_life != 16) %>% 
    dplyr::select(gene_id, gene_name, group, Half_life)
) %>% mutate(group = factor(group, 
                            levels = c("Soma, mESC", "Neurite, mESC", "Nucleus, PCN", "Soma, PCN", "Neurite, PCN")
                            )
             ) -> tmp.ridges

ggplot(data = tmp.ridges, aes(y = group, x = Half_life, fill = group)) + 
  geom_density_ridges(
    scale = .95, rel_min_height = .01
  ) +
  scale_fill_manual(name = "Compartment", 
                    values = alpha(unname(c(pal_loc.2way, pal_loc.3way)[c(3, 1, 7, 5, 4)]), 0.67)) + 
  scale_x_continuous(breaks = seq(0, 16, by = 4), limits = c(0, 16)) +
  labs(y = "", x = "mRNA Half-life, hours") +
  theme_bw(base_size = 14) -> p.2.1

ggsave("../output/p.2.1.pdf")

## statistical test --- Tukey
# TukeyHSD(aov(formula = Half_life ~ group, data = p.2.1$data))$group %>% 
#   as_tibble(., 
#             rownames = "comparison") %>% 
#             mutate(test = "Tukey post-hoc test for difference in means",
#                    group = "", 
#                    variable = "",
#                    value = "Half-life",
#                    `p-value` = "") %>% 
#   dplyr::select(group, variable, value, comparison, test, `p-value`, `p-adjusted` = `p adj`) %>% 
#   mutate(sign.symb = case_when(
#            `p-adjusted` < 0.001 ~ "***",
#            `p-adjusted` < 0.01 ~ "**",
#            `p-adjusted` < 0.05 ~ "*",
#            T ~ "NS"
#          )) -> p.2.1_stats
# write_tsv(p.2.1_stats, 
#           "../output/p.2.1_stats.tsv")
## Update 06.06.2023 ---- Use pairwise comparisons with Welch's t-test
p.2.1$data %>% 
  mutate(group = as.character(group),
         ranks = rank(Half_life, ties.method = "average"),
         lg2trsfrm = log2(Half_life)) -> dat.trsfrmd

purrr::map_dfr(unique(dat.trsfrmd$group), 
               function(ref) {
                 comps <- as.character(unique(dat.trsfrmd$group))[unique(dat.trsfrmd$group) != ref]
                 purrr::map_dfr(comps,   
                                function(comp) {
                                  dat.tmp <- dat.trsfrmd %>% filter(group %in% c(ref, comp))
                                  rbind(
                                    tibble(comparison = paste0(ref, " vs ", comp),
                                           data.type = "raw", test = "Welch's t-test", 
                                           p.value = t.test(Half_life ~ group, data = dat.tmp)$p.value),
                                    tibble(comparison = paste0(ref, " vs ", comp),
                                           data.type = "log2-transformed", test = "Welch's t-test", 
                                           p.value = t.test(lg2trsfrm ~ group, data = dat.tmp)$p.value),
                                    tibble(comparison = paste0(ref, " vs ", comp),
                                           data.type = "rank-transformed", test = "Welch's t-test", 
                                           p.value = t.test(ranks ~ group, data = dat.tmp)$p.value)
                                  )
                                })
               }) %>% 
  mutate(dupli = duplicated(p.value)) %>% 
  filter(!dupli) %>% 
  group_by(data.type) %>% 
  mutate(p.adjusted = p.adjust(p.value, "fdr")) %>% 
  dplyr::select(-dupli) -> t.pvals

write_tsv(t.pvals, 
          "../output/p.2.1_stats.tsv")
```


################################################################################
## 2.2 Probabilistic grid showing the chance of an mRNA with an average Half-life within a given compartment to reach neuritic distance X.
```{r}
res_phen.long_3c <- read_rds("../data/PCN.ThreeWay_dnCaf1.CompSep/DEres.3c_Phen.Long.RDS") %>% 
  mutate(comp_1 = str_split_fixed(comp, "_", 2)[, 1],
         comp = str_remove_all(comp, "Pro_|Cyto_|Nuc_"))
res_comp.long_3c <- read_rds("../data/PCN.ThreeWay_dnCaf1.CompSep/DEres.3c_Comp.Long.RDS") %>% 
  inner_join(mm10Ann, ., by = c("gene_name")) %>% 
  filter(str_count(comp, "GFP") == 2 | str_count(comp, "GFP") == 0) %>% 
  mutate(
    comp_1 = str_split_fixed(comp, "_", 3)[, 3],
    comp = str_remove_all(comp, "_GFP|_dnCaf1")
    )

l.hl.comps$PCN.3c %>% 
               mutate(sign = case_when(
                        group == "Nuc" ~ "Nucleus",
                        group == "Cyto" ~ "Cytoplasm",
                        group == "Pro" ~ "Projections",
                        T ~ "ayy"
                      ),
                      gene_id = str_split_fixed(gene, "_", 2)[, 1],
                      gene_name = str_split_fixed(gene, "_", 2)[, 2])  %>% 
  inner_join(.,
             rbind(
               res_comp.long_3c %>% 
                 filter(comp_1 == "GFP", comp == "Pro.vs.Cyto", sign != "NS") %>% 
                 dplyr::select(gene_name, sign),
               res_comp.long_3c %>% 
                 filter(comp_1 == "GFP", comp == "Cyto.vs.Nuc", sign == "Nucleus") %>% 
                 dplyr::select(gene_name, sign)
             ),
             by = c("gene_name", "sign")) -> tab.3c 
tab.3c %>% 
  group_by(sign) %>% 
  summarise(n = n(),
            mean = mean(Half_life),
            median = median(Half_life), .groups = "drop") -> stat.3c

read_tsv("../data/Zappulo_2017/Zappulo_soft-filt_Proc_vs_Soma_deseq_res.tsv",
         show_col_types = F) %>% 
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
         gene_name = str_split_fixed(gene, "_", 2)[, 2],
         sign = case_when(
    log2FoldChange > 0.58 & padj < 0.05 ~ "Neurites",
    log2FoldChange < -0.58 & padj < 0.05 ~ "Soma",
    T ~ "NS"
  )) %>% 
  dplyr::select(gene_id, gene_name, sign) %>% 
  inner_join(., l.hl.comps$Ascl1.2c, by = c("gene_id", "gene_name", "sign" = "Compartment")) -> tab.2c 
tab.2c %>% 
  group_by(sign) %>% 
  summarise(n = n(),
            mean = mean(Half_life),
            median = median(Half_life), .groups = "drop") -> stat.2c
stat.loc <- rbind(stat.2c, stat.3c)
##
stat.loc %>% 
  dplyr::rename(localization = sign, n_genes = n, 
                `mean Half-life, hours` = mean,
                `median Half-life, hours` = median) %>% 
  write_tsv(., "../output/p.2.2_summary.tsv")
```


```{r}
dendprob_bg <- purrr::map2(1:100, 
                              seq(0.1, 16, length.out = 100), 
            ~ list(name = .x, 
                   tr_hl = .y, 
                   nbins = 100, 
                   ddist = 500, 
                   DR = 3.0 * 10^(-3), 
                   uR = 5.8 * 10^(-3), 
                   betaR = 0.0109) %>% 
              rdd()
            ) %>% 
  purrr::reduce(., rbind)

ggplot(data = dendprob_bg, aes(x = dend_dist, y = rna_hl)) + 
  geom_raster(aes(fill = rna_pdens)) + 
  geom_hline(data = stat.loc %>% mutate(Localization = factor(sign,
                                                              levels = c("Neurite",
                                                                         "Soma",
                                                                         "Projections",
                                                                         "Cytoplasm",
                                                                         "Nucleus"))), 
             aes(yintercept = mean, 
                 color = Localization)) +
  scico::scale_fill_scico(palette = "lajolla", begin = 0, name = "Probability:") +
  scale_color_manual(values = c(pal_loc.2way, pal_loc.3way)[c(1, 3:5, 7)]) +
  scale_y_continuous(breaks = seq(0, 16, by = 4)) +
  labs(y = "mRNA half-life, hours", x = "Neurite distance, mm") +
  theme(aspect.ratio = 1, 
        panel.grid = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(size = 0.25),
        axis.ticks = element_line(size = 0.25)) -> p.2.2

ggsave(p.2.2, filename = "../output/p.2.2.pdf", width = 4.5, height = 4.5)
```


################################################################################
## 2.3 Probability to venture further than 100 micrometers of Neuritically & Somatically localized transcripts from 3c PCN data.
```{r}
tab.3c %>% 
  filter(sign != "Nucleus") %>% 
  dplyr::select(gene_name, Half_life, sign) %>% 
  split(., .$sign) %>% 
  purrr::map2_dfr(.,
                  names(.),
                  ~ purrr::map2_dfr(.x$gene_name,
                                    .x$Half_life,
                                    function(gene, hl) {
                                      list(name = gene, 
                                           tr_hl = hl, 
                                           nbins = 5, 
                                           ddist = 100, 
                                           DR = 3.0 * 10^(-3), 
                                           uR = 5.8 * 10^(-3), 
                                           betaR = 0.0109) %>% 
                                      rdd()
                                    }) %>% mutate(sign = .y)
                  ) -> t.p_1fd

ggplot(data = t.p_1fd %>% filter(dend_dist == 100), 
       aes(x = rna_pdens)) +
  geom_histogram(aes(fill = sign),
                 color = "black", 
                 size = 0.1, 
                 bins = 18, position = "identity") + 
  geom_vline(xintercept = 0.5, color = "red", size = 0.5) +
  scale_fill_manual(values = alpha(pal_loc.3way[1:2], 0.67)) + 
  coord_cartesian(xlim = c(0, 1)) +
  theme(aspect.ratio = 1) +
  labs(y = "Frequency",
       x = "Probability to reach >= 100 microm", 
       fill = "Localization") -> p.2.3

ggsave(p.2.3, filename = "../output/p.2.3.pdf", width = 4.5, height = 3.5)


tibble(comparison = "Projection vs Cytoplasm", 
       test = "Welch's", 
       p.value = t.test(rna_pdens ~ sign, data = p.2.3$data)$p.value) -> p.2.3_stats
write_tsv(p.2.3_stats, "../output/p.2.3_stats.tsv")
```


########################################################  
#   3. Figure 4 plots
################################################################################
## 3.1 Histograms showing the distributions of m6A-enriched & m6A-depleted transcripts along lg2FC between subcellular compartments
RNA-sequencing data from both 3- and 2- way compartment separation were used.
```{r}
read_tsv("../data/hard-filt_outfilt_IP_unfr_m6a.sysy_vs_FLOW_unfr_m6a.sysy_deseq_res.tsv", 
         show_col_types = F) %>% 
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
         gene_name = str_split_fixed(gene, "_", 2)[, 2]) %>% 
  dplyr::select(-gene) -> res_msa.sysy
## filter by lfcSE
res_msa.sysy <- res_msa.sysy %>% 
  filter(lfcSE < quantile(res_msa.sysy$lfcSE, 0.9))


### Load 3c data from PCN
loc.3c <- read_rds("../data/PCN.ThreeWay_dnCaf1.CompSep/DEres.3c_Comp.Wide.RDS")
loc.3c <- loc.3c %>%
filter(comp == "Cyto.vs.Pro",
         lfcSE_GFP < quantile(loc.3c$lfcSE_GFP, 0.9))

### Load 2c data from mESC
loc.mesc <- read_tsv("../data/Zappulo_2017/Zappulo_soft-filt_Proc_vs_Soma_deseq_res.tsv",
                     show_col_types = F)
loc.mesc <- loc.mesc %>%
  filter(lfcSE < quantile(loc.mesc$lfcSE, 0.9)) %>% 
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
         gene_name = str_split_fixed(gene, "_", 2)[, 2]) %>% 
  dplyr::select(-gene) %>% 
  dplyr::select(contains("gene_"), everything())

##### Assemble the plot for 3c
res_msa.sysy %>% 
  mutate(
    msa_var = case_when(
      log2FoldChange > 1 & padj < 0.05 ~ "m6A-enriched",
      log2FoldChange < -1 & padj < 0.05 ~ "m6A-depleted",
      T ~ "NS"
    )
  ) %>% 
  filter(msa_var != "NS") %>% 
  dplyr::select(contains("gene_"), msa_var) %>% 
  inner_join(.,
             loc.3c,
             by = c("gene_name")) -> t_p.3.1.3c

ggplot(data = t_p.3.1.3c, aes(x = log2FoldChange_GFP)) + 
  geom_histogram(aes(fill = msa_var),
                 position = "identity", 
                 bins = 50,
                 color = "black", 
                 size = 0.25,
                 alpha = 0.5) +
  geom_vline(xintercept = 0, color = "red") +
  scale_fill_manual(values = pal_msa[c(1,3)]) +
  labs(x = "", y = "Frequency", fill = "m6A status") -> p.3.1_3c


tibble(
  group = "localisation 3c",
  comparison = "m6A-enriched.vs.m6A-depleted",
  test = "Welch's two sample t-test",
  `p-value` = t.test(formula = log2FoldChange_GFP ~ msa_var, data = t_p.3.1.3c)$p.value,
) %>% 
  mutate(`p-adjusted` = p.adjust(`p-value`, method = "fdr"),
         sign.symb = case_when(
            `p-adjusted` < 0.001 ~ "***",
            `p-adjusted` < 0.01 ~ "**",
            `p-adjusted` < 0.05 ~ "*",
            T ~ "NS"
            )
         ) -> p.3.1.3c_stats
write_tsv(p.3.1.3c_stats, "../output/p.3.1.3c_stats.tsv")
ggsave(plot = p.3.1_3c, filename = "../output/p.3.1.3c.pdf", width = 6, height = 4)

##### Assemble the plot for 2c
res_msa.sysy %>% 
  mutate(
    msa_var = case_when(
      log2FoldChange > 1 & padj < 0.05 ~ "m6A-enriched",
      log2FoldChange < -1 & padj < 0.05 ~ "m6A-depleted",
      T ~ "NS"
    )
  ) %>% 
  filter(msa_var != "NS") %>% 
  dplyr::select(contains("gene_"), msa_var) %>% 
  inner_join(.,
             loc.mesc,
             by = c("gene_id", "gene_name")) -> t_p.3.1.2c

ggplot(data = t_p.3.1.2c, aes(x = log2FoldChange)) + 
  geom_histogram(aes(fill = msa_var),
                 position = "identity", 
                 bins = 50,
                 color = "black", 
                 size = 0.25,
                 alpha = 0.5) +
  geom_vline(xintercept = 0, color = "red") +
  scale_fill_manual(values = pal_msa[c(1,3)]) +
  scale_x_continuous(breaks = seq(-6, 6, by = 2)) +
  labs(x = "", y = "Frequency", fill = "m6A status") -> p.3.1.2c


tibble(
  group = "localisation 2c (mESC)",
  comparison = "m6A-enriched.vs.m6A-depleted",
  test = "Welch's two sample t-test",
  `p-value` = t.test(formula = log2FoldChange ~ msa_var, data = t_p.3.1.2c)$p.value,
) %>% 
  mutate(`p-adjusted` = p.adjust(`p-value`, method = "fdr"),
         sign.symb = case_when(
            `p-adjusted` < 0.001 ~ "***",
            `p-adjusted` < 0.01 ~ "**",
            `p-adjusted` < 0.05 ~ "*",
            T ~ "NS"
            )
         ) -> p.3.1.2c_stats
write_tsv(p.3.1.2c_stats, "../output/p.3.1.2c_stats.tsv")
ggsave(plot = p.3.1.2c, filename = "../output/p.3.1.2c.pdf", width = 6, height = 4)
```

################################################################################
## 3.2 Box-plots showing change in expression and localisation of m6A-enriched/-depleted transcripts in shYTHDF and shMETTL3 KD
```{r}
read_rds("../data/PCN.TwoWay_shYthdf123/DEres.2c_Phen.Long.RDS") -> res_phen.shyth
read_rds("../data/PCN.TwoWay_shMettl3/DEres.2c_Phen.Long.RDS") -> res_phen.shmet
msa_tab.cons <- read_tsv("/local/artem/Projects/Stability/Results/MultipleDatasetsConsensus/consensus_msa_tab.tsv", 
                    show_col_types = F)

list(
  res_phen.shyth,
  res_phen.shmet
) %>% 
  purrr::map_dfr(.,
                 ~ .x %>% mutate(comp_1 = str_split_fixed(comp, "_", 2)[, 1],
                                 comp = str_remove_all(comp, "Neurite_|Soma_"),
                                 comp = str_remove_all(comp, ".vs.Scrambled")) %>% 
                   inner_join(msa_tab.cons[msa_tab.cons$dataset_msa == "consensus", ], .,
                              by = c("gene_id" = "gene_name"))
                 ) %>% 
  mutate(msa_var = factor(msa_var, levels = c("m6A depleted", "Not significant", "m6A enriched"))) -> t_p.3.2.Dexp

t_p.3.2.Dexp %>% 
  group_by(gene_id, msa_var, comp) %>% 
  summarise(log2FoldChange = mean(log2FoldChange), .groups = "drop") %>% 
ggplot(data = .,
       aes(x = comp, y = log2FoldChange)) + 
  geom_boxplot(aes(fill = msa_var), outlier.shape = NA, size = 0.25) + 
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = alpha(pal_msa, 0.67), name = "m6A") +
  #facet_grid(. ~ comp_1) + 
  coord_cartesian(ylim = c(-2, 2)) + 
  theme(aspect.ratio = 2) -> p.3.2.Dexp

ggsave("../output/p.3.2.Dexp.pdf", p.3.2.Dexp)

t_p.3.2.Dexp %>% 
  group_by(gene_id, msa_var, comp) %>% 
  summarise(log2FoldChange = mean(log2FoldChange), .groups = "drop") %>% 
  filter(msa_var != "Not significant") %>% 
  group_by(comp) %>% 
  nest() %>% 
  mutate(test = "Welch's two sample t-test",
         comparison = "Upregulation.vs.m6A-enrichment",
         `p-value` = purrr::map(data, ~ t.test(formula = log2FoldChange ~ msa_var, data = .x)$p.value)) %>% 
  unnest(`p-value`) %>% 
  ungroup() %>% 
  mutate(`p-adjusted` = p.adjust(`p-value`, method = "fdr"),
         sign.symb = case_when(
            `p-adjusted` < 0.001 ~ "***",
            `p-adjusted` < 0.01 ~ "**",
            `p-adjusted` < 0.05 ~ "*",
            T ~ "NS"
            )
         ) %>% 
  dplyr::select(group = comp, test, comparison, `p-value`, `p-adjusted`, sign.symb) -> p.3.2.Dexp_stats
write_tsv(p.3.2.Dexp_stats, "../output/p.3.2.Dexp_stats.tsv")
```



```{r}
# read_rds("../data/PCN.TwoWay_shYthdf123/DEres.2c_Comp.Wide.RDS") %>% 
#   mutate(delta_loc = log2FoldChange_shYTHDF123 - log2FoldChange_Scrambled,
#          comp_1 = "shYTHDF123") -> res_comp.shyth
# 
# read_rds("../data/PCN.TwoWay_shMettl3/DEres.2c_Comp.Wide.RDS") %>% 
#   mutate(delta_loc = log2FoldChange_shMettl3 - log2FoldChange_Scrambled,
#          comp_1 = "shMettl3") -> res_comp.shmet
# 
# 
# list(
#   res_comp.shyth %>% dplyr::select(gene_id = gene_name, delta_loc, comp_1),
#   res_comp.shmet %>% dplyr::select(gene_id = gene_name, delta_loc, comp_1)
# ) %>% 
#   purrr::map_dfr(.,
#                  ~ .x %>% 
#                    inner_join(msa_tab.cons[msa_tab.cons$dataset_msa == "consensus", ], .,
#                               by = c("gene_id"))
#                  ) %>% 
#   mutate(msa_var = factor(msa_var, levels = c("m6A depleted", "Not significant", "m6A enriched"))) -> t_p.f5h
```



################################################################################
## 3.3 Box-plots showing expression and change in localisation of selected GO-terms in shYTHDF PCNs
```{r}
read_rds("../data/PCN.TwoWay_shYthdf123/DEres.2c_Comp.Wide.RDS") %>% 
  inner_join(mm10Ann, ., by = c("gene_id" = "gene_name")) %>% 
  mutate(delta_loc = log2FoldChange_shYTHDF123 - log2FoldChange_Scrambled,
         comp_1 = "shYTHDF123") -> res_comp
# Load in DGEs from shYTHDF run
read_rds("../data/PCN.TwoWay_shYthdf123/DEres.2c_Phen.Long.RDS") %>% 
  inner_join(mm10Ann, ., by = c("gene_id" = "gene_name")) -> res_phen
dds <- read_rds("../data/PCN.TwoWay_shYthdf123/DDS.2c.RDS")
normTransform(dds) %>% 
  assay() %>% 
  as_tibble(., rownames = "gene_id") %>% 
  pivot_longer(., !contains("gene"), names_to = "group", values_to = "cts.norm") %>% 
  mutate(comp = str_split_fixed(group, "_", 3)[, 2],
         group = str_split_fixed(group, "_", 3)[, 1]) %>% 
  group_by(gene_id, group, comp) %>% 
  summarise(cts.avg = mean(cts.norm), .groups = "drop") %>% 
  inner_join(mm10Ann, ., by = "gene_id") -> cts

res_phen %>% 
  filter(lfcSE < quantile(res_phen$lfcSE, 0.75),
         sign != "NS") %>% 
  mutate(comp_1 = str_split_fixed(comp, "_", 2)[, 1]) -> res_phen.tmp

up.shyth <- res_phen.tmp %>% 
  filter(sign == "shYTHDF123") %>% 
  split(., .$comp_1) %>% 
  purrr::map(., ~ .x %>% pull(gene_name)) %>% 
  purrr::reduce(., intersect)
up.shyth_n <- res_phen.tmp %>% 
  filter(sign == "shYTHDF123", comp_1 == "Neurite") %>% 
  pull(gene_name)


# Prepare a GO annotation file
mgi <- read_tsv("../data/auxiliary/gene_association.mgi", skip = 3, col_names = F)
mgi.ann <- read_tsv("../data/auxiliary/go_terms.mgi", skip = 3, col_names = F)
## prepare separate gene annotation file
mgi %>% 
  dplyr::select(X3, X12, X10) %>% 
  distinct(X3, X10, .keep_all = T) %>% 
  `colnames<-`(c("gene_name", "gene_biotype", "info")) %>% 
  distinct(gene_name, .keep_all = T) -> gene.ann

## we actually can drop categorical terms to clean up a lot of space
mgi %>% 
  filter(!str_detect(X4, "NOT")) %>% 
  dplyr::select(gene_name = X3, GO.term = X5) %>% 
  inner_join(., 
             mgi.ann %>% dplyr::select(GO.term = X2, GO.term_category = X1, GO.term_name = X3),
             by = "GO.term") %>% 
  filter(!GO.term_name %in% c("molecular_function", "biological_process", "cellular_component")) %>% 
  distinct(gene_name, GO.term, GO.term_category, GO.term_name) -> go2gene.ann
  
## Append additional terms that were absent in the inital table from MGI
suppGO.n2n <- read_tsv("../data/PCN.TwoWay_shYthdf123/GO_term_summary_n2n.txt",
                       show_col_types = F) %>% 
  distinct(Symbol, .keep_all = T) %>% 
  mutate(GO.term = "GO:0098984",
         GO.term_category = "Cellular Component",
         GO.term_name = "neuron to neuron synapse") %>% 
  dplyr::select(gene_name = Symbol, contains("GO.term"))
suppGO.mit <- read_tsv("../data/PCN.TwoWay_shYthdf123/GO_term_summary_mito.txt", 
                       show_col_types = F) %>% 
  distinct(Symbol, .keep_all = T) %>% 
  mutate(GO.term = "GO:0098798",
         GO.term_category = "Cellular Component",
         GO.term_name = "mitochondrial protein−containing complex") %>% 
  dplyr::select(gene_name = Symbol, contains("GO.term"))
##
go2gene.ann <- list(go2gene.ann, suppGO.n2n, suppGO.mit) %>% purrr::reduce(., rbind)


## Load canonical terms related to stable and localized transcripts identified in PCNs before
gost.stloc <- read_tsv("../data/GO_terms.clpsd.tsv",
                        show_col_types = F)
t_gos.stloc <- go2gene.ann %>% 
  filter(GO.term %in% gost.stloc$term_id)

#### Assemble the plot
t_gos.stloc_upn <- t_gos.stloc %>% 
  filter(gene_name %in% up.shyth_n)
## Keep only terms with > 15 genes
t_gos.stloc_upn %>% 
  tabyl(GO.term_name) %>% 
  as_tibble() %>% 
  filter(n > 15) %>% 
  pull(GO.term_name) -> gos.tokeep.n


## Gene expression plot
t_gos.stloc_upn %>% 
  inner_join(cts, ., by = c("gene_name")) %>% 
  filter(GO.term_name %in% gos.tokeep.n,
         comp == "Neurite") %>% 
  ggplot(data = ., aes(x = cts.avg, y = str_wrap(GO.term_name, 42))) + 
  geom_boxplot(aes(fill = group), size = 0.25, outlier.size = 0.5, outlier.alpha = 0.3) + 
  #geom_vline(xintercept = cts_summary %>% filter(str_detect(dummy, "Neurite")) %>% pull(cts.mean), size = 0.5) +
  scale_fill_manual(values = alpha(pal_mut.msa[c(1, 3)], 0.67)) +
  theme(aspect.ratio = 0.75, legend.position = "bottom") + 
  labs(x = "log2(CPM + 1)", y = "GO term", fill = "") -> p_stLocGOs.shYthdf123UpN_GExpr

t_gos.stloc_upn %>% 
  inner_join(res_comp, ., by = c("gene_name")) %>% 
  filter(GO.term_name %in% gos.tokeep.n) %>% 
  filter(log2FoldChange_shYTHDF123 < quantile(res_comp$log2FoldChange_shYTHDF123, 0.75),
         log2FoldChange_Scrambled < quantile(res_comp$log2FoldChange_Scrambled, 0.75)) %>% 
  mutate(delta_loc = log2FoldChange_shYTHDF123 - log2FoldChange_Scrambled) %>% 
  ggplot(data = ., aes(x = delta_loc, y = GO.term_name)) + 
  geom_boxplot(size = 0.25, outlier.size = 0.5, outlier.alpha = 0.3) + 
  geom_vline(xintercept = 0, size = 0.5) +
  theme(aspect.ratio = 0.35) + 
  labs(x = "delta lg2FC Neurite vs Soma;\nshYthdf123 - Scrambled",
       y = "GO term") -> p_stLocGOs.shYthdf123UpN_deltaLoc
list(
  p_stLocGOs.shYthdf123UpN_GExpr,
  p_stLocGOs.shYthdf123UpN_deltaLoc
) %>% purrr::map2(.,
                  c("../output/p.3.3.shYthdf123UpN_GExpr",
                    "../output/p.3.3.shYthdf123UpN_deltaLoc"),
                   function(x, xname) {
                     ggsave(plot = x, filename = paste0("../output/", xname, ".pdf"),
                            width = 7, height = 4.5, device = "pdf")
                   })
```


################################################################################
## 3.4 Analysis of electrophysiological data from shYTHDF PCNs
```{r}
t_spikes <- read_tsv("../data/PCN.TwoWay_shYthdf123/t_spikes_yth.tsv", show_col_types = F)

## filter inactive wells cells in D3 and D2 die at day 21, only residual activity at day 25
t.yth <- t_spikes %>% 
  filter(!(Well %in% c("D3", "D2") & DIV == "25"),
         Treatment %in% c("Scrambled", "ShYTFD"), 
         !Well %in% c("D5"))

## summarise per DIV for line plotting
t.yth_sum <- t.yth %>%  
  group_by(Treatment, DIV) %>% 
  summarise(MFR_br.avg = mean(MFR, na.rm = T),
            MFR_br.sd = sd(MFR, na.rm = T),
            .groups = "drop") %>% 
  mutate(DIV = as.numeric(DIV))

#### Assemble the plot
update_geom_defaults("point", list(colour = NULL))
ggplot(data = t.yth %>% 
         mutate(DIV = factor(DIV, levels = c(7, 11, 14, 18, 21, 25))), 
       aes(x = DIV, y = MFR)) + 
  geom_boxplot(aes(fill = Treatment), size = 0.25, outlier.shape = 21) + 
  scale_fill_manual(values = alpha(pal_mut.msa[c(1,3)], 0.67)) + 
  theme(aspect.ratio = 1.25) + 
  labs(y = "MFR (Hz / nAE)") -> p.mfr_boxes
## revert to default options for durther plotting
update_geom_defaults("point", list(colour = "black"))

ggsave(filename = "../output/p.3.4.pdf", p.mfr_boxes)


### Raster snap shot of activity
l.div_mhz <- read_rds("../data/PCN.TwoWay_shYthdf123/l.assembled_yth.RDS")
t.meta <- read_tsv("../data/PCN.TwoWay_shYthdf123/t_meta.tsv", show_col_types = F)
## I will look specifically at DIV18 as it has the largest differences in spike features
div18_mhz <- l.div_mhz$`18`
## Only keep wells that are relevant to the analysis
div18_mhz <- div18_mhz[t.yth[t.yth$DIV == "18", ]$Well]

purrr::map2_dfr(div18_mhz,
            names(div18_mhz),
            function(fr_list, well) {
              fr_list$firing_rates$frate %>% 
                as_tibble(., rownames = "bin") %>% 
                pivot_longer(., !contains("bin"), names_to = "Electrode", values_to = "spikes") %>% 
                mutate(Well = well) %>% 
                dplyr::select(Well, Electrode, bin, spikes)
            }) %>% 
  inner_join(t.meta, ., by = "Well") -> t.raster_mhz

#### Visualise
t.raster_mhz_sniped <- t.raster_mhz %>% 
  filter(Well %in% c("D2", "B4")) %>% 
  mutate(bin = factor(bin, levels = 0:9999))


ggplot(data = t.raster_mhz_sniped, 
       aes(y = Electrode, x = bin)) + 
  geom_raster(aes(fill = spikes), interpolate = F) +
  #geom_vline(xintercept = c(200, 280)) +
  #facet_grid(Treatment + Well ~ ., scales = "free") + 
  facet_wrap(Treatment ~ .,
             scales = "free_y",
             nrow = 2,
             strip.position = "left",
             labeller = as_labeller(c(Scrambled = "Scrambled\nElectrodes",
                                      ShYTFD = "shYthdf123\nElectrodes")))  +
  scale_fill_gradient2(low = "white", high = "black") + 
  scale_x_discrete(breaks = seq(0, 10000, by = 1000), 
                   labels = seq(0, 100, by = 10)) +
  coord_cartesian(xlim = c(0, 9000)) +
  theme(axis.text.y = element_blank(),
        #axis.ticks.y = element_blank(), 
        axis.line.y = element_blank(),
        panel.border = element_rect(size = 0.25, fill = NA)) +
  labs(y = NULL, x = "Recording time, s") +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "none",
        aspect.ratio = 0.125) -> p.raster_toshow
ggsave(plot = p.raster_toshow, 
       filename = paste0("../output/p.3.4.raster.pdf"),
       width = 6,
       height = 2.5)
```


### Permutation test ----
To test unambigously test for the differences between case/control we perform a MWU-test 5000 times
on the same data with randomly permuted labels
```{r}
## Auxiliary function to perform a permutation test
perm_test <- function(tab, col_to_perm, col_to_test, n_perm = 5000) {
  grps <- unique(tab[[col_to_perm]])
  purrr::map_dfr(1:n_perm,
                 function(iter) {
                   tab[[col_to_perm]] <- sample(grps, 
                                                dim(tab)[1],
                                                replace = T)
                   
                   tibble(n_iter = iter,
                          p.value = wilcox.test(formula = !!sym(col_to_test) ~ !!sym(col_to_perm), 
                                                data = tab)$p.value)
                 })
}

t.yth %>%
  filter(DIV %in% c("14", "18", "21", "25")) %>% 
  wilcox.test(formula = MFR ~ Treatment, data = .) -> spks.test_mwu
t.yth %>%
  filter(DIV %in% c("14", "18", "21", "25")) %>% 
  perm_test(tab = ., col_to_perm = "Treatment", col_to_test = "MFR", n_perm = 5000) -> t.perm
t.perm <- t.perm %>% 
  mutate(m.lg10_pval = -log10(p.value))
tibble(comparison = "shYTHDF vs Scrambled", test = "Permutation",
       p.value = diff(ecdf(t.perm$m.lg10_pval)(c(-log10(spks.test_mwu$p.value), Inf)))) -> p.3.4_stats

write_tsv(p.3.4_stats, "../output/p.3.4_stats.tsv")
```


########################################################  
#   4. Figure 5 plots
################################################################################
## 4.1 MA-plots in shnELAVL for Neurites and Soma. Downregulated transcripts are coloured in red (lg2FC >= log2(1.5))
```{r}
dds.shelavl <- read_rds("../data/PCN.TwoWay_shElavl234/DDS.2c.RDS")
purrr::map2_dfr(c("Soma_shElavl234", "Neurite_shElavl234"),
                c("Soma_Scrambled", "Neurite_Scrambled"),
                ~ results(dds.shelavl, contrast = c("condition", .x, .y)) %>% 
                  as_tibble(., rownames = "gene_id") %>% 
                  mutate(comp = paste0(.x, ".vs.", .y))) %>% 
  inner_join(mm10Ann, ., by = "gene_id") -> res_long.shelavl

res.shelavl_n <- res_long.shelavl %>% 
  drop_na() %>% 
  filter(comp == "Neurite_shElavl234.vs.Neurite_Scrambled") %>% 
  mutate(
    sign = case_when(
      log2FoldChange < -log2(1.5) & padj < 0.05 ~ "Downregulated",
      T ~ "NS"
      ),
    sign = ifelse(lfcSE > 1, "NS", sign)
    )

ggplot(data = res.shelavl_n, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  #geom_point(data = res.pvsc_slam %>% filter(gene_name %in% l.mrks$Psynapse), color = "red", size = 0.75) +
  # ggrepel::geom_label_repel(data = res.shelavl_n %>% filter(gene_name %in% l.mrks$Psynapse), 
  #                           aes(label = gene_name),
  #                           force = 92, nudge_x = 2, nudge_y = 1) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = alpha(c("red", "gray50"), 0.67), name = "shElavl234") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.4.1.Neurite

res.shelavl_s <- res_long.shelavl %>% 
  drop_na() %>% 
  filter(comp == "Soma_shElavl234.vs.Soma_Scrambled") %>% 
  mutate(
    sign = case_when(
      log2FoldChange < -log2(1.5) & padj < 0.05 ~ "Downregulated",
      T ~ "NS"
      ),
    sign = ifelse(lfcSE > 1, "NS", sign)
    )

ggplot(data = res.shelavl_s, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  #geom_point(data = res.pvsc_slam %>% filter(gene_name %in% l.mrks$Psynapse), color = "red", size = 0.75) +
  # ggrepel::geom_label_repel(data = res.shelavl_n %>% filter(gene_name %in% l.mrks$Psynapse), 
  #                           aes(label = gene_name),
  #                           force = 92, nudge_x = 2, nudge_y = 1) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = alpha(c("red", "gray50"), 0.67), name = "shElavl234") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.4.1.Soma

ggsave(p.4.1.Neurite, filename = "../output/p.4.1.Neurite.pdf",
       width = 5, height = 5)
ggsave(p.4.1.Soma, filename = "../output/p.4.1.Soma.pdf",
       width = 5, height = 5)
```


################################################################################
## 4.2 Boxplots showing the distributions of AREs (in UTR) of transcripts downregulated by ELAVls KD.
```{r}
##
res_comp.wide_shelavl <- read_rds("../data/PCN.TwoWay_shElavl234/DEres.2c_Comp.Wide.RDS") %>% 
  inner_join(mm10Ann, ., by = c("gene_id" = "gene_name"))
v.shelavl_down <- read_gmt("../data/PCN.TwoWay_shElavl234/shElavl234.vs.Scrambled_N&S_lg2fc0.58.gmt")$Scrambled
t.are <- read_tsv("../data/TPU_ARE.cts.tsv", show_col_types = F)


##
rbind(
  res_comp.wide_shelavl %>% 
    filter(gene_name %in% v.shelavl_down) %>% 
    mutate(cat = "shElavl234 Down S+N"),
  res_comp.wide_shelavl %>% 
    mutate(cat = "Population")
) %>% 
  mutate(delta_loc = log2FoldChange_shElavl234 - log2FoldChange_Scrambled) %>% 
  inner_join(., t.are[, c("gene_id", "TPU.length", "ARE.cts")], by = "gene_id") -> t.shelavl

##
ggplot(data = t.shelavl, aes(x = ARE.cts, y = cat)) + 
  geom_boxplot(aes(fill = cat),
               size = 0.25, 
               color = "black", 
               outlier.colour = "gray50",
               outlier.alpha = 0.35,
               outlier.size = 0.75) +
  scale_fill_manual(values = c("gray50", "red")) +
  coord_cartesian(xlim = c(0, 40)) + 
  labs(y = "", x = "ARE count") +
  theme(aspect.ratio = 0.35, legend.position = "none") -> p.4.2

ggsave(plot = p.4.2, filename = "../output/p.4.2.pdf", width = 5, height = 5.5)

rbind(
  p.4.2$data %>% filter(str_detect(cat, "Elavl")),
  p.4.2$data %>% filter(!str_detect(cat, "Elavl")) %>% slice_sample(n = 336) ## number of transcripts in a group of interest
) -> tmp

tibble(
group = "",
variable = "",
value = "ARE count",
comparison = "shElavl234 Down S+N.vs.Random sample",
test = "Welch's two sample t-test",
`p-value` = t.test(formula = ARE.cts ~ cat, data = tmp)$p.value
) %>% 
  mutate(`p-adjusted` = p.adjust(`p-value`, method = "fdr"),
         sign.symb = case_when(
         `p-adjusted` < 0.001 ~ "***",
         `p-adjusted` < 0.01 ~ "**",
         `p-adjusted` < 0.05 ~ "*",
         T ~ "NS")
         ) -> p.4.2_stats
write_tsv(p.4.2_stats, "../output/p.4.2_stats.tsv")
```



################################################################################
## 4.3. Boxplots showing the change in localisation for targets of ELAVls (transcripts that are down-regulated upon ELAVls KD).
```{r}
ggplot(data = t.shelavl, aes(x = delta_loc, y = cat)) + 
  geom_boxplot(aes(fill = cat),
               size = 0.25, 
               color = "black", 
               outlier.colour = "gray50",
               outlier.alpha = 0.35,
               outlier.size = 0.75) +
  scale_fill_manual(values = c("gray50", "red")) +
  geom_vline(xintercept = 0, size = 0.5) +
  scale_x_continuous(breaks = seq(-6, 6, by = 1)) +
  labs(x = "delta lg2FC Neurite vs Soma\nshElavl234 - Scrambled", y = "") + 
  theme(aspect.ratio = 0.35, legend.position = "none") -> p.4.3

ggsave(plot = p.4.3, filename = "../output/p.4.3.pdf", width = 6, height = 3)

rbind(
  t.shelavl %>% filter(str_detect(cat, "Elavl")),
  t.shelavl %>% filter(!str_detect(cat, "Elavl")) %>% slice_sample(n = 336)
) -> tmp

tibble(
group = "",
variable = "",
value = "delta lg2FC shElavl234 - Scrambed",
comparison = "shElavl234 Down S+N.vs.Random sample",
test = "Welch's two sample t-test",
`p-value` = t.test(formula = delta_loc ~ cat, data = tmp)$p.value
) %>% 
  mutate(`p-adjusted` = p.adjust(`p-value`, method = "fdr"),
         sign.symb = case_when(
         `p-adjusted` < 0.001 ~ "***",
         `p-adjusted` < 0.01 ~ "**",
         `p-adjusted` < 0.05 ~ "*",
         T ~ "NS")
         ) -> p.4.3_stats
write_tsv(p.4.3_stats, "../output/p.4.3_stats.tsv")
```


################################################################################
## 4.4. Bar plot showing the GO terms overrepresented among targets of ELAVls (transcripts that are down-regulated upon ELAVls KD).
```{r}
gost(list(shElavl234 = v.shelavl_down), organism = "mmusculus", ordered_query = F, significant = T, evcodes = T) -> gostres

gostres$result %>% 
  dplyr::select(-evidence_codes) %>% 
  split(., .$query) %>% 
  purrr::map_dfr(., 
                 function(gores) {
                   gores %>%
                     #dplyr::select() %>% 
                     filter(str_detect(source, "GO|KEGG"), 
                            term_size < 500,
                            !str_detect(term_name, "Corona|Hippo")) %>% 
                     arrange(term_size) -> gores.filt
                   purrr::map_dfr(gores.filt$term_id,
                                  function(term) {
                                    gores.filt %>% filter(term_id == term) -> gores.term
                                    gores.term %>% pull(parents) %>% unlist() -> v.pars
                                    ## break and return if the parent of the starting term is not present
                                    if (!any(v.pars %in% gores.filt$term_id)) {
                                        return(
                                          gores.term %>% 
                                            mutate(parents = paste0(unlist(parents), collapse = ","))
                                        )
                                      }
                                    v.chlds <- gores.term$term_id
                                    ## look along parents until there are no more detected terms
                                    while (length(v.pars) > 0) {
                                      v.pars.out <- v.pars
                                      gores.filt %>% filter(term_id %in% v.pars) %>% pull(parents) %>% unlist() -> v.pars
                                      if (!any(v.pars %in% gores.filt$term_id)) {
                                        break
                                      }
                                      v.chlds <- c(v.chlds, v.pars)
                                    }
                                    ## remove the last element of the child vector since it the term we want to keep
                                    v.chlds <- v.chlds[!(v.chlds %in% v.pars.out)] ## not used after all
                                    ## update the table
                                    gores.filt %>% 
                                      filter(term_id %in% v.pars.out) %>% 
                                      mutate(parents = paste0(unlist(parents), collapse = ","))
                                  }
                                  ) %>% distinct(term_id, .keep_all = T)
                 } 
                 ) -> t.gostres_clpsd

goa_mouse <- read_tsv("../data/auxiliary/goa_mouse.gpi", skip = 22, col_names = F, show_col_types = F)
goa_mouse <- goa_mouse %>% dplyr::select(gene_name = X3, info = X4)

terms <- unique(t.gostres_clpsd$term_id)
terms.names <- t.gostres_clpsd[which(terms %in% t.gostres_clpsd$term_id), ]$term_name
purrr::map2_dfr(terms, terms.names,
                function(term, name) {
                  t.gostres_clpsd %>% filter(term_id == term) %>% dplyr::slice(1) %>% pull(intersection) %>% 
                    str_split_fixed(., ",", Inf) %>% 
                    as.character() -> term.genes
                  goa_mouse %>% filter(gene_name %in% term.genes) %>% 
                    mutate(GO.term_id = term,
                           GO.term_name = name) %>% 
                    dplyr::select(contains("GO"), everything())
                }) -> t.gostres_genes

write_tsv(t.gostres_genes, "../output/p.4.4_GO.termAnnotation.tsv")

split(t.gostres_clpsd, t.gostres_clpsd$query) %>% 
  purrr::map(.,
             ~ .x %>% 
               mutate(term_name = str_wrap(term_name, 42)) %>% 
               group_by(source) %>% 
               arrange(p_value) %>% 
               dplyr::slice(1:5) %>% 
               #arrange(desc(intersection_size / query_size)) %>% 
               #filter(source != "TF") %>% 
               ggplot(data = ., aes(x = 100 * intersection_size / query_size, y = term_name)) + 
               geom_col(aes(fill = -log10(p_value)), size = 0.25, color = "black") + 
               facet_grid(source ~ ., scales = "free") + 
               scale_fill_gradientn(colors = RColorBrewer::brewer.pal(9, "Reds"), limits = c(0, 10)) + 
               labs(x = "Intersection / query, %", y = "Term name")) -> l.p_GOshelavl

ggsave(plot = l.p_GOshelavl$shElavl234, filename = "../output/p.4.4.pdf", width = 6, height = 4)
```


########################################################  
#   5. Figure 6 plots
################################################################################
## 5.1 ECDFs showing the the distributions of localisation of transcripts with the most optimal (green) and suboptimal codon composition (blue)
```{r}
## results of the ORFEOME library sequencing
res_orfm <- read_tsv("../data/t.orfm.res.tsv", show_col_types = F)

ggsave(plot = plot_tai_ecdf(res_orfm, 0.1), 
       filename = "../output/p.5.1.pdf",
       width = 4, height = 3.75)
```

########################################################  
#   6. Plots related to Supplementary Figure 1
################################################################################
## 6.1. MA-plots for compartment separation RNA-seq runs including: 3c-PCN Slamseq WT, 3c-PCN RNA-seq GFP + dnCAF1, 2c-Ascl1 Slamseq WT
```{r}
l.mrks <- list()
l.mrks[["Soma"]] <- c("Snhg11", "Ncoa3", "Malat1", "Miat", "Son")
l.mrks[["Nucleus"]] <- c("Hnrnph1","Snhg11", "Miat", "Malat1", "Meg3")
l.mrks[["Psynapse"]] <- c("Homer2", "Map2", "Cox5b", "Kif5a", "Camk2a")#, "Gap43", "Sptbn4", "Nav1", "Ank3", "Nefh", "Nefl", "Mapt")
l.mrks[["Soma.2c"]] <- c("Nup210", "Snhg11", "Xist")
l.mrks[["Neurite.2c"]] <- c("Tagln", "Col3a1", "Lamc1", "Lamb1", "Lama1", "Myo1c")


dds.sl3c <- read_rds("../data/PCN.ThreeWay_Slamseq.Rerun/DDS.3cSlam.RDS")
purrr::map2_dfr(c("Pro", "Cyto"),
                c("Cyto", "Nuc"),
                ~ results(dds.sl3c, contrast = c("group", .x, .y)) %>% 
                  as_tibble(., rownames = "gene_id") %>% 
                  mutate(comp = paste0(.x, ".vs.", .y))) %>% 
  inner_join(mm10Ann, ., by = "gene_id") -> res_sl3c.long

res.pvsc_slam <- res_sl3c.long %>% 
  drop_na() %>% 
  filter(comp == "Pro.vs.Cyto") %>% 
  mutate(
    sign = case_when(
      log2FoldChange > log2(1.5) & padj < 0.05 ~ "Projections",
      log2FoldChange < -log2(1.5) & padj < 0.05 ~ "Cytoplasm",
      T ~ "NS"
      ),
    sign = ifelse(lfcSE > 1, "NS", sign)
    )

ggplot(data = res.pvsc_slam, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  geom_point(data = res.pvsc_slam %>% filter(gene_name %in% l.mrks$Psynapse), color = "red", size = 0.75) +
  ggrepel::geom_label_repel(data = res.pvsc_slam %>% filter(gene_name %in% l.mrks$Psynapse), 
                            aes(label = gene_name),
                            force = 92, nudge_x = 2, nudge_y = 1) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = pal_loc.3way[1:3], name = "Localization") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.sf1_MA.3c.Slamseq_Pro.vs.Cyto

res.cvsn_slam <- res_sl3c.long %>% 
  drop_na() %>% 
  filter(comp == "Cyto.vs.Nuc") %>% 
  mutate(
    sign = case_when(
      log2FoldChange > log2(1.5) & padj < 0.05 ~ "Cytoplasm",
      log2FoldChange < -log2(1.5) & padj < 0.05 ~ "Nucleus",
      T ~ "NS"
      ),
    sign = ifelse(lfcSE > 1, "NS", sign)
    )

ggplot(data = res.cvsn_slam, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  geom_point(data = res.cvsn_slam %>% filter(gene_name %in% l.mrks$Nucleus), color = "red", size = 0.75) +
  ggrepel::geom_label_repel(data = res.cvsn_slam %>% filter(gene_name %in% l.mrks$Nucleus), 
                            aes(label = gene_name),
                            force = 177) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = pal_loc.3way[2:4], name = "Localization") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.sf1_MA.3c.Slamseq_Cyto.vs.Nuc


####
res_3c.long <- read_rds("../data/PCN.ThreeWay_dnCaf1.CompSep/DEres.3c_Comp.Long.RDS")
res.pvsc_gfp <- res_3c.long %>% 
  filter(comp == "Pro_GFP.vs.Cyto_GFP") %>% 
  mutate(comp = str_remove_all(comp, "_GFP")) %>% 
  inner_join(mm10Ann[, c("gene_id", "gene_name", "description")], ., by = "gene_name")

ggplot(data = res.pvsc_gfp, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  geom_point(data = res.pvsc_gfp %>% filter(gene_name %in% l.mrks$Psynapse), color = "red", size = 0.75) +
  ggrepel::geom_label_repel(data = res.pvsc_gfp %>% filter(gene_name %in% l.mrks$Psynapse), 
                            aes(label = gene_name),
                            force = 92, nudge_x = 2, nudge_y = 1) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = pal_loc.3way[1:3], name = "Localization") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.sf1_MA.3c.GFP_Pro.vs.Cyto

res.cvsn_gfp <- res_3c.long %>% 
  filter(comp == "Cyto_GFP.vs.Nuc_GFP") %>% 
  mutate(comp = str_remove_all(comp, "_GFP")) %>% 
  inner_join(mm10Ann[, c("gene_id", "gene_name", "description")], ., by = "gene_name")

ggplot(data = res.cvsn_gfp, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  geom_point(data = res.cvsn_gfp %>% filter(gene_name %in% l.mrks$Nucleus), color = "red", size = 0.75) +
  ggrepel::geom_label_repel(data = res.cvsn_gfp %>% 
                              filter(gene_name %in% l.mrks$Nucleus), 
                            aes(label = gene_name),
                            force = 269, nudge_y = -1.5) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = pal_loc.3way[2:4], name = "Localization") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.sf1_MA.3c.GFP_Cyto.vs.Nuc


res.pvsc_dncaf1 <- res_3c.long %>% 
  filter(comp == "Pro_dnCaf1.vs.Cyto_dnCaf1") %>% 
  mutate(comp = str_remove_all(comp, "_GFP|_dnCaf1")) %>% 
  inner_join(mm10Ann[, c("gene_id", "gene_name", "description")], ., by = "gene_name")

ggplot(data = res.pvsc_dncaf1, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  geom_point(data = res.pvsc_dncaf1 %>% filter(gene_name %in% l.mrks$Psynapse), color = "red", size = 0.75) +
  ggrepel::geom_label_repel(data = res.pvsc_dncaf1 %>% filter(gene_name %in% l.mrks$Psynapse), 
                            aes(label = gene_name),
                            force = 92, nudge_x = 2, nudge_y = 1) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = pal_loc.3way[1:3], name = "Localization") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.sf1_MA.3c.dnCAF1_Pro.vs.Cyto

res.cvsn_dncaf1 <- res_3c.long %>% 
  filter(comp == "Cyto_dnCaf1.vs.Nuc_dnCaf1") %>% 
  mutate(comp = str_remove_all(comp, "_GFP|_dnCaf1")) %>% 
  inner_join(mm10Ann[, c("gene_id", "gene_name", "description")], ., by = "gene_name")

ggplot(data = res.cvsn_dncaf1, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  geom_point(data = res.cvsn_dncaf1 %>% filter(gene_name %in% l.mrks$Nucleus), color = "red", size = 0.75) +
  ggrepel::geom_label_repel(data = res.cvsn_dncaf1 %>% filter(gene_name %in% l.mrks$Nucleus), 
                            aes(label = gene_name),
                            force = 269, nudge_y = -1.51) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = pal_loc.3way[2:4], name = "Localization") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.sf1_MA.3c.dnCAF1_Cyto.vs.Nuc

## 2c Slamseq Ascl1
dds.2cslam <- read_rds("../data/Ascl1.Compartments_Slamseq/DDS.Slam.RDS")

purrr::map2_dfr(c("Soma_shElavl234", "Neurite_shElavl234"),
                c("Soma_Scrambled", "Neurite_Scrambled"),
                ~ results(dds.shelavl, contrast = c("condition", .x, .y)) %>% 
                  as_tibble(., rownames = "gene_id") %>% 
                  mutate(comp = paste0(.x, ".vs.", .y))) %>% 
  inner_join(mm10Ann, ., by = "gene_id") -> res_long.shelavl
results(dds.2cslam, contrast = c("group", "Soma", "Neurite")) %>% as_tibble(., rownames = "gene_id") -> tmp

ggplot(data = tmp, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point() + 
  ylim(c(-5, 5))

res_2c <- read_tsv("../data/Ascl1.Compartments_Slamseq/DE.tab_Tall.tsv")
res.nvss_2c.slam <- res_2c %>% 
  inner_join(mm10Ann[, c("gene_id", "gene_name", "description")], ., by = c("gene_id" = "gene")) %>% 
  dplyr::rename(sign = loc.var)

ggplot(data = res.nvss_2c.slam, aes(x = log2(baseMean), y = log2FoldChange)) + 
  geom_point(aes(color = sign, alpha = sign), size = 0.35) +
  geom_point(data = res.nvss_2c.slam %>% filter(gene_name %in% c(l.mrks$Soma.2c, l.mrks$Neurite.2c)), 
             color = "red", 
             size = 0.75) +
  ggrepel::geom_label_repel(data = res.nvss_2c.slam %>% filter(gene_name %in% c(l.mrks$Soma.2c, l.mrks$Neurite.2c)), 
                            aes(label = gene_name),
                            force = 169, nudge_y = .51) +
  geom_hline(yintercept = c(-0.58, 0, 0.58), linetype = c("dashed", "solid", "dashed")) +
  scale_color_manual(values = pal_loc.2way[1:3], name = "Localization") +
  scale_alpha_manual(values = c(0.7, 0.2, 0.7)) +
  guides(color = guide_legend(override.aes = list(size = 3)),
         alpha = "none") +
  scale_y_continuous(breaks = seq(-6, 6, by = 2)) +
  theme(aspect.ratio = 1) -> p.sf1_MA.2c_N.vs.S


##
purrr::map2(list(p.sf1_MA.3c.GFP_Cyto.vs.Nuc,
                 p.sf1_MA.3c.GFP_Pro.vs.Cyto,
                 p.sf1_MA.3c.dnCAF1_Cyto.vs.Nuc,
                 p.sf1_MA.3c.dnCAF1_Pro.vs.Cyto,
                 p.sf1_MA.3c.Slamseq_Pro.vs.Cyto,
                 p.sf1_MA.3c.Slamseq_Cyto.vs.Nuc,
                 p.sf1_MA.2c_N.vs.S),
            c("p.6.1_MA.3c.GFP_Cyto.vs.Nuc",
              "p.6.1_MA.3c.GFP_Pro.vs.Cyto",
              "p.6.1_MA.3c.dnCAF1_Cyto.vs.Nuc",
              "p.6.1_MA.3c.dnCAF1_Pro.vs.Cyto",
              "p.6.1_MA.3c.Slamseq_Pro.vs.Cyto",
              "p.6.1_MA.3c.Slamseq_Cyto.vs.Nuc",
              "p.6.1_MA.2c_N.vs.S"),
            function(MAplot, plotname) {
              ggsave(plot = MAplot, filename = paste0("../output/", plotname, ".pdf"), width = 5, height = 5)
            })
```


################################################################################
## 6.2. Boxplots showing T to C conversion rates estimated from Slamseq runs
```{r}
## aux
plot_ct <- function(tab_t0nn, ylims = c(-2, 10)) {
  ggplot(data = tab_t0nn, 
       aes(x = conv, y = CR.avg * 100, color = col_var)) + 
    geom_boxplot(outlier.shape = NA) + 
    scale_color_manual(values = c("grey30", "red"), guide = "none") +
    stat_summary(fun.data = function(x) tibble(y = -(0.25 * 8), 
                                               label = round(mean(x), 2)), 
                 geom = "text", 
                 size = 4.5,
                 vjust = 0.5, 
                 hjust = 0.1, 
                 angle = 90) +
    #geom_hline(yintercept = 0) +
    coord_cartesian(ylim = ylims) +
    scale_y_continuous(breaks = seq(0, 10, by = 2)) +
    #facet_grid(. ~ group) +
    #ggforce::facet_zoom(ylim = c(0, 0.6), zoom.size = , 
    #zoom.data = ifelse(conversion != "T->C", NA, FALSE), horizontal = FALSE) +
    theme(aspect.ratio = 1,
          axis.text.x = element_text(angle = 45, 
                                     hjust = 1, 
                                     colour = c(rep("black", 10), "red", "black"))
    ) +
    labs(y = "Conversion rate (%)", 
         x = "")
}

### mESC neurons
tab_cr_T0NN.mESC <- read_tsv("../data/cr_T0.NN_mESC.tsv",
                            show_col_types = F)
plot_ct(tab_cr_T0NN.mESC, 
        ylims = c(-2, 6.5)
        ) -> p.6.2.mesc

### Whole Primary cortical neurons
tab_cr_T0NN.PCN <- read_tsv("../data/cr_T0.NN_PCN.tsv",
                            show_col_types = F)
plot_ct(tab_cr_T0NN.PCN,
        ylims = c(-2, 8)
        ) -> p.6.2.pcnwhole

### 3-way compartment separation of Primary cortical neurons
tab_cr_T0NN.PCN3c <- read_tsv("../data/cr_T0.NN_PCN3c.tsv",
                              show_col_types = F)
plot_ct(tab_cr_T0NN.PCN3c, 
        ylims = c(-2, 6)
        ) -> p.6.2.pcn3c

ggsave(plot = p.6.2.mesc, filename =  paste0("../output/p.6.2.mESC.pdf"), width = 4, height = 4)
ggsave(plot = p.6.2.pcnwhole, filename =  paste0("../output/p.6.2.PCN.pdf"), width = 4, height = 4)
ggsave(plot = p.6.2.pcn3c, filename =  paste0("../output/p.6.2.PCN3c.pdf"), width = 4, height = 4)
```

################################################################################
## 6.3. Boxplots showing normalized conversion rates per replicate along timepoints in Slamseq experiments.
```{r}
t.cr.n <- read_tsv("../data/PCN.ThreeWay_Slamseq.Rerun/PCN3c_CR.TC_filt.norm.tsv",
                   show_col_types = F) %>%  
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
         gene_name = str_split_fixed(gene, "_", 2)[, 2],
         group = case_when(
           group == "Pro" ~ "Neurite",
           group == "Nuc" ~ "Nucleus",
           group == "Cyto" ~ "Cytoplasm",
           T ~ "ayy"
         ),
         group = factor(group, levels = c("Nucleus", "Cytoplasm", "Neurite"))) %>% 
  filter(gene_id %in% unique(str_split_fixed(l.hl.comps$PCN.3c$gene, "_", 2)[, 1])) %>% 
  dplyr::select(contains("gene_"), group, timepoint, b.rep, cr_raw, cr_norm)

t.cr.n %>% 
  filter(timepoint != 0) %>% 
  group_by(gene_id, gene_name, group, timepoint) %>% 
  summarise(cr_norm.avg = mean(cr_norm),
            cr_norm.sd = sd(cr_norm), .groups = "drop") %>% 
  mutate(timepoint = factor(timepoint, levels = c("4", "8", "16"))) %>% 
  ggplot(data = , aes(x = timepoint, y = cr_norm.avg)) + 
  geom_boxplot(aes(fill = group), outlier.shape = NA, size = 0.25) +
  scale_fill_manual(values = alpha(unname(pal_loc.3way)[c(4,2,1)], 0.67)) + 
  coord_cartesian(ylim = c(0, 1)) +
  guides(colour = "none") + 
  labs(y = "Normalized T->C conversion rate", x = "Timepoint, hours", fill = "Compartment") + 
  theme(aspect.ratio = 1, legend.position = c(0.85, 0.8)) -> p.6.3


ggsave(p.6.3, filename = "../output/p.6.3.pdf", width = 5.5, height = 5.5, units = "in")
```


################################################################################
## 6.4. 2-dimensional density plot showing correlation of delta Half-live and localisation of mRNAs between Neurites and Soma for 2c and 3c compartment separations.
```{r}
### 3c
hlcomps.3c <- tmp.ridges %>%
  filter(str_detect(group, "PCN"), !str_detect(group, "Nucleus")) %>% 
  mutate(group = str_remove(group, ", PCN")) %>% 
  pivot_wider(names_from = "group", values_from = "Half_life") %>% 
  mutate(delta_hl = Neurite - Soma)
res.pvsc_gfp %>% 
  inner_join(., hlcomps.3c, by = c("gene_id", "gene_name")) %>%
  #filter(enr.var != "NS") %>% 
  #filter(padj < 0.05) %>% 
  ggplot(data = ., aes(x = log2FoldChange, y = delta_hl)) + 
  geom_bin_2d(color = "gray50", size = 0.1, bins = 60) +
  scale_fill_gradientn(colours = RColorBrewer::brewer.pal(9, name = "Oranges")[2:9]) +
  geom_vline(xintercept = 0, size = 0.5) + 
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(method = "lm", formula = y ~ x) +
  #geom_density_2d() +
  #geom_point(alpha = 0.15, size = 0.3) +
  ggpubr::stat_cor(size = 4.5, label.y = 15, label.x = -5) + 
  coord_cartesian(x = c(-5, 5), y = c(-9, 16)) +
  scale_x_continuous(breaks = seq(-4, 4, by = 2)) +
  scale_y_continuous(breaks = seq(-8, 16, by = 4)) +
  labs(x = "lg2FC Neurite vs Soma", y = "delta mRNA Half-life, hours\nNeurite vs Soma") + 
  theme(aspect.ratio = 1,
        legend.position = c(0.1, 0.65),
        legend.background = element_blank()) -> p.6.4.3c

ggsave(plot = p.6.4.3c, 
       filename = "../output/p.6.4.3c.pdf", 
       width = 5, height = 5)


### 2c
hlcomps.2c <- read_tsv("../data/Ascl1.Compartments_Slamseq/data_hl.softfilt.wide.tsv", show_col_types = F) %>% 
  drop_na() %>% 
  mutate(delta_hl = Neurite - Soma)

res.zap <- read_tsv("/local/artem/Projects/Stability/Code/data/Zappulo_2017/Zappulo_soft-filt_Proc_vs_Soma_deseq_res.tsv",
                    show_col_types = F)
res.zap %>% 
  mutate(gene_id = str_split_fixed(gene, "_", 2)[, 1],
         gene_name = str_split_fixed(gene, "_", 2)[, 2]) %>% 
  dplyr::select(-gene) %>% 
  dplyr::select(contains("gene"), everything()) %>% 
  inner_join(., hlcomps.2c, by = c("gene_id", "gene_name")) %>%
  ggplot(data = ., aes(x = log2FoldChange, y = delta_hl)) + 
  geom_bin_2d(color = "gray50", size = 0.1, bins = 60) +
  scale_fill_gradientn(colours = RColorBrewer::brewer.pal(9, name = "Oranges")[2:9]) +
  geom_vline(xintercept = 0, size = 0.5) + 
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(method = "lm", formula = y ~ x) +
  ggpubr::stat_cor(size = 4.5, label.y = 15, label.x = -5) + 
  coord_cartesian(x = c(-5, 5), y = c(-9, 16)) +
  scale_x_continuous(breaks = seq(-4, 4, by = 2)) +
  scale_y_continuous(breaks = seq(-8, 16, by = 4)) +
  labs(x = "lg2FC Neurite vs Soma", y = "delta mRNA Half-life, hours\nNeurite vs Soma") + 
  theme(aspect.ratio = 1,
        legend.position = c(0.1, 0.65),
        legend.background = element_blank()) -> p.6.4.2c

ggsave(plot = p.6.4.2c, 
       filename = "../output/p.6.4.2c.pdf", 
       width = 5, height = 5)
```


########################################################  
#   7. Plots related to Supplementary Figure 5
################################################################################
## 7.1. Bar-plots showing fractions of m6A- enriched/depleted transcripts among transcripts that are up-/down- regualted upon shMettl3 and shYthdf123 KDs.
```{r}
de.shmettl3 <- read_gmt("../data/PCN.TwoWay_shMettl3/shMettl3.vs.Scrambled_N&S_lg2fc0.58.gmt")
de.shythdf <- read_gmt("../data/PCN.TwoWay_shYthdf123/shYthdf123.vs.Scrambled_N&S_lg2fc0.58.gmt")
msa_tab.cons %>% 
  inner_join(mm10Ann, ., by = "gene_id") %>% 
  mutate(shMettl3 = case_when(
    gene_name %in% de.shmettl3$shMettl3 ~ "Up-regulated",
    gene_name %in% de.shmettl3$Scrambled ~ "Down-regulated",
    gene_name %in% de.shmettl3$NS ~ "NS",
    T ~ "NA"
  )) %>% 
  mutate(shYthdf123 = case_when(
    gene_id %in% de.shythdf$shYTHDF123 ~ "Up-regulated",
    gene_id %in% de.shythdf$Scrambled ~ "Down-regulated",
    gene_id %in% de.shythdf$NS ~ "NS",
    T ~ "NA"
  )) -> tab.plot

ggplot(data = tab.plot %>% filter(shYthdf123 != "NA"), aes(x = shYthdf123, fill = msa_var)) + 
  geom_bar(position = "fill", size = 0.25, color = "black") +
  scale_fill_manual(values = alpha(pal_msa[c(3, 1, 2)], 0.67), name = "m6A consensus:") +
  theme(aspect.ratio = 1.5) -> p.7.1.shythdf
ggplot(data = tab.plot %>% filter(shMettl3 != "NA"),
       aes(x = shMettl3, fill = msa_var)) + 
  geom_bar(position = "fill", size = 0.25, color = "black") +
  scale_fill_manual(values = alpha(pal_msa[c(3, 1, 2)], 0.67), name = "m6A consensus:") +
  theme(aspect.ratio = 1.5) -> p.7.1.shmettl3

ggsave(p.7.1.shythdf, filename = "../output/p.7.1.shythdf.pdf")
ggsave(p.7.1.shmettl3, filename = "../output/p.7.1.shmettl3.pdf")



tab.plot %>% 
  filter(shYthdf123 != "Not significant", shYthdf123 != "NA") %>% 
  tabyl(shYthdf123, msa_var) %>% 
  chisq.test() -> p.7.1.shythdf.tst
tab.plot %>% 
  filter(shMettl3 != "Not significant", shMettl3 != "NA") %>% 
  tabyl(shMettl3, msa_var) %>% 
  chisq.test() -> p.7.1.shmettl3.tst
tibble(Group = c("shYthdf123", "shMettl3"), 
       comparison = "m6A status vs DE", 
       test = "chi-squared", 
       p.value = c(p.7.1.shythdf.tst$p.value, 
                   p.7.1.shmettl3.tst$p.value)) %>% 
  mutate(p.adj = p.adjust(p.value)) %>% 
  write_tsv(., "../output/p.7.1_stats.tsv")
```

################################################################################
## 7.1. Changes in localisation between neurites and soma for DE transcripts in shMettl3 and shYthdf123 KDs.
```{r}
de.shythdf_wide <- read_rds("../data/PCN.TwoWay_shYthdf123/DEres.2c_Comp.Wide.RDS")
de.shmettl3_wide <- read_rds("../data/PCN.TwoWay_shMettl3/DEres.2c_Comp.Wide.RDS")
##

de.shythdf_wide  %>% 
  inner_join(mm10Ann, ., by = c("gene_id" = "gene_name")) %>%
  mutate(shYthdf123 = case_when(
    gene_id %in% de.shythdf$shYTHDF123 ~ "Up-regulated",
    gene_id %in% de.shythdf$Scrambled ~ "Down-regulated",
    gene_id %in% de.shythdf$NS ~ "NS",
    T ~ "NA"
  )) %>% 
  mutate(delta_loc = log2FoldChange_shYTHDF123 - log2FoldChange_Scrambled) -> dloc.shythdf
  
de.shmettl3_wide %>% 
  inner_join(mm10Ann, ., by = c("gene_id" = "gene_name")) %>%
  mutate(shMettl3 = case_when(
    gene_name %in% de.shmettl3$shMettl3 ~ "Up-regulated",
    gene_name %in% de.shmettl3$Scrambled ~ "Down-regulated",
    gene_name %in% de.shmettl3$NS ~ "NS",
    T ~ "NA"
  )) %>% 
  mutate(delta_loc = log2FoldChange_shMettl3 - log2FoldChange_Scrambled) -> dloc.shmettl3

###
ggplot(data = dloc.shythdf %>% filter(!shYthdf123 %in% c("NA", "NS")),
       aes(x = delta_loc)) +
  geom_density(aes(fill = shYthdf123), size = 0.25, color = "black") + 
  geom_vline(xintercept = 0, size = 0.5) +
  scale_fill_manual(values = alpha(c("gray50", "red"), 0.37)) +
  scale_x_continuous(breaks = seq(-2, 2, by = 1), limits = c(-2, 2)) + 
  labs(y = "", x = "delta lg2FC Neurite vs Soma\nshYthdf123 - Scrambled") -> p.7.2.shythdf

ggplot(data = dloc.shmettl3 %>% filter(!shMettl3 %in% c("NA", "NS")),
       aes(x = delta_loc)) +
  geom_density(aes(fill = shMettl3), size = 0.25, color = "black") + 
  geom_vline(xintercept = 0, size = 0.5) +
  scale_fill_manual(values = alpha(c("gray50", "red"), 0.37)) +
  scale_x_continuous(breaks = seq(-2, 2, by = 1), limits = c(-2, 2)) + 
  labs(y = "", x = "delta lg2FC Neurite vs Soma\nshMettl3 - Scrambled") -> p.7.2.shmettl3

ggsave(p.7.2.shythdf, filename = "../output/p.7.2.shythdf.pdf")
ggsave(p.7.2.shmettl3, filename = "../output/p.7.2.shmettl3.pdf")

ks.test(formula = delta_loc ~ shYthdf123, data = p.7.2.shythdf$data) -> p.7.2.shythdf.tst
ks.test(formula = delta_loc ~ shMettl3, data = p.7.2.shmettl3$data) -> p.7.2.shmettl3.tst
tibble(Group = c("shYthdf123", "shMettl3"), 
       comparison = "m6A status vs DE", 
       test = "chi-squared", 
       p.value = c(p.7.2.shythdf.tst$p.value, 
                   p.7.2.shmettl3.tst$p.value)) %>% 
  mutate(p.adj = p.adjust(p.value)) %>% 
  write_tsv(., "../output/p.7.2_stats.tsv")
```

########################################################  
#   8. Plots related to Supplementary Figure 7
################################################################################
## 8.1. Probabilities of a transcripts to reach a dendritic distance of 400 micrometers with/without the help of active transport (based on a model from Fonkeu et al. 2019)
```{r}
t.dist <- purrr::pmap(list(
  rep(c(0.5, 6, 27), 2),
  rep(c(0, 1), each = 3),
  rep(c("diffusion only", "with active transport"), each = 3)
  ),
            ~ list(name = ..1, 
                   tr_hl = ..1, 
                   nbins = 100, 
                   ddist = 400, 
                   DR = 3.0 * 10^(-3), 
                   uR = ..2 * 5.8 * 10^(-3), 
                   betaR = 0.0109) %>% 
              rdd() %>% mutate(group = ..3)
            ) %>% 
  purrr::reduce(., rbind)

ggplot(data = t.dist, aes(x = dend_dist, y = rna_pdens)) + 
  geom_line(aes(color = group)) + 
  facet_grid(. ~ rna_hl) + 
  labs(x = "Neuritic distance, micrometers", y = "Probability", color = "") +
  theme(aspect.ratio = 0.75) -> p.8.1

ggsave(p.8.1, filename = "../output/p.8.1.pdf")
```



